---
title: "Assignment 1 - Intro to R (10 points)"
author: "CARSON CRENSHAW (cgc8gdt)"
output: html_document
---

For this assignment, please use "<-" as the assignment operator rather than "=".

## Question 1 (1 point)
>Write an equation with multiple terms and assign it to `x` so `x` will be equal to `11`
```{r}
# Following PEMDAS, the equation with multiple terms below will result in 11. The variable 'x' will be written as to illustrate the correct result in the console (print out the value of the variable 'x'). 

x <- ((20*13) - (149-10))/11
x

```

## Question 2 (1 point)
>Write an equation with multiple terms and save it to `y` so `y` will be equal to `7`
```{r}
# Following PEMDAS, the equation with multiple terms below will result in 7. The variable 'y' will be written as to illustrate the correct result in the console (print out the value of the variable 'y'). 

y <- ((13*14)/7) - 19
y

```

## Question 3 (1 point)
> Write an equation between `x` and `y` that gives an output of `100`
*you can be as straight forward or creative as you want (remember PEMDAS)*
```{r}
# The equation below using both terms 'x' and 'y' as previously defined will result in the numerical value 100. 

(x*11) - (y*3)

```

## Question 4 (1 point)
> Determine the class of 'v' with a function call (v is defined within code block)
```{r}
# Using the function call class, the variable 'v' has a numeric class. This means that the values within the vector are numeric. 

v <- c(25, 11, 6, 27, 4, 2)
class(v)

```

## Question 5 (1 point)
> Select the number `6` from the vector `v` by numerical index
```{r}
# The number '6' from the vector 'v' is selected using a numerical index. Square brackets are used to select the third element out of the vector.

v[3]

```

## Question 6 (1 point)
> Select the 4th element in the vector `v`
```{r}
# The 4th element in the vector 'v' can be selected using a numerical index as well. This results in the value 
# 27. 

v[4]

```

## Question 7 (1 point)
> Select the first 3 elements of the vector `v` using a vector index (`c()`)
```{r}
# The first three elements of the vector 'v' found using a vector index are 25, 11, and 6. 

v[c(1, 2, 3)]


```

## Question 8 (1 point)
>Select the first 3 elements of the vector `v` using a `:` for the indexing
```{r}
# The first three elements of the vector 'v' found using a 'vector index':' for the indexing are also 25, 11, and 6. 

v[1:3]

```

## Question 9 (1 point)
> Select the first 2 elements and the last element of the vector `v`
```{r}
# The first 2 elements of the vector 'v' are 25 and 11, and the last element is 2. 

v[c(1:2, 6)]

```

## Question 10 (1 point)
> Produce a binary vector where TRUE denotes elements greater than 5 in the vector v
```{r}
# To produce a binary vector where TRUE denotes elements greater than 5 in the vector v, one must use logical operators to compare elements in a vector to a number. Four values in the vector are TRUE (greater than 5), and two are FALSE.  

v > 5

# One can then define the operation above to create a new vector.

new_v <- v > 5
new_v

```








