---
title: "Assignment 3 - Descriptive Stats (20 pts)"
author: "CARSON CRENSHAW (cgc8gdt)"

output: html_document
---

**You will submit the html output and raw rmd file.**

This assignment will use the popular Titanic dataset.
Find the data dictionary for the dataset here: 
https://www.kaggle.com/competitions/titanic/data
Download the file "train.csv" from the kaggle site, and rename the file to "titanic.csv"

```{r}
# Load the dataset; assign the dataframe to object called 'titanic'

titanic <- read.csv('titanic.csv')

```

## Question 1 (.5 point)
```{r}
# Look at the first 6 rows of data

head(titanic)

```

## Question 2 (2 point)
```{r}
# sort rows of titanic datafame by fare (smallest to largest) and assign to "sorted_titanic"

sorted_titanic <- titanic[order(titanic$Fare), ]
sorted_titanic

```

## Question 3 (1 point)
```{r}
# subset the dataframe to include passengers in 2nd class and include the columns for class, name, age, and sex

subset(titanic, (Pclass == 2), c(Pclass, Name, Age, Sex))

```

## Question 4 (2 point)
```{r}
# How many NAs are in the full dataset? (show with code) 

# This code ensures that both empty numeric and character entries are represented as NA
titanic[titanic==""] <- NA
sum(is.na(titanic))


# How many NAs are in the 'age' column? (hint: you may need to Google) (show with code)
sum(is.na(titanic$Age))


```

## Question 5 (1 point)
```{r}
# Recode the variable 'sex' of the dataset as factor so that 'male' becomes 'm' and 'female' becomes 'f'

# Coding the variable Sex as a factor variable allows the user to change the name of the levels by using the levels()
factor_titanic <- factor(titanic$Sex)
levels(factor_titanic) <- c("f", "m")
titanic$Sex = factor_titanic
titanic

```

## Question 6 (.5 point)
```{r}
# print summary statistics for the dataset

# The summary() function prints the five number summary + the mean for each column
summary(titanic)

```

## Question 7 (1 point)
```{r}
# What is the average ticket price for a first class ticket?
# (round answer to 2 decimal points)

# In order to calculate the average ticket price (Fare) for a first class (Pclass = 1) ticket, one must subset the data and then find the mean of the Fare column
round(colMeans(subset(titanic, (Pclass == 1), Fare)), 2)

```

## Question 8 (1 point)
```{r}
# Plot a histogram for the fare price

# Histogram showing all fare prices for the titanic dataset
hist(titanic$Fare,xlab = 'Fare Price', main = 'Price Distribution', xlim = c(0,600), ylim = c(0, 800))

```

## Question 9 (1 point)
```{r}
# Using 'histogram' function in lattice package (required for question) to plot a histogram of fare price 
# separated by port of embarkation (embarked)
# If unfamiliar, you will need to look up the lattice package
# histogram(~ variable | dividing variable, data = data_name)

# Importing the lattice package necessary to run the code
# Histogram split between the three embarkation ports. Because the two empty values (both at $80) were marked as NA using earlier code, these values do not appear in this histogram. 
library('lattice')
histogram(~Fare|Embarked,data=titanic,
       xlab="Fare Price",
       main="Titanic Fare Price by Port of Embarkation")

```


## Question 10 (1 point)
```{r}
# Write one sentence interpreting this graph

#There was a more spread out distribution of fares for people who embarked at Cherbourg (C) compared to Southampton (S) and Queenstown (Q), meaning that a variety of passangers across classes embarked at Cherbourg whereas it was far more common to only have passangers with less expensive/lower class tickets embark from the other ports. 

```

## Question 11 (1 point)
```{r}
# Create a barplot of counts of cabin class vs sex
# x axis is sex, bars are shaded by cabin class (use one of the commands for 
# contingency table)

barplot(table(titanic$Pclass,titanic$Sex),
        beside = TRUE,
        xlab = "Sex (Male or Female)",
        ylab = "Number of People in a Cabin Class",
        ylim = c(0, 350),
        legend.text = c("Class1", "Class2","Class3"),
        main = 'Cabin Class vs. Sex')

```

## Question 12 (4 point)
```{r}
# (1) create a vector of 10 numbers which includes a few NAs 
# (2) confirm it is a vector with a built-in function
# (3) sum the number of NAs in vector
# (4) remove the NA values from your vector and then delete your vector 
# object from the workspace

#1 
practice_vector <- c(1,2,3,NA,5,6,7,NA,9,NA)
#2
is.vector(practice_vector)
#3
sum(is.na(practice_vector))
#4
practice_vector[!is.na(practice_vector)]
rm(practice_vector) 
# the remove() function would also work here to clear the current R working environment of this object

```

## Question 13 (4 points)
```{r}
# (1) create a matrix of numbers from 1 to 500 going by 2's that is 25 rows by 10 columns 
# (2) confirm it is a matrix using a built-in function
# (3) get the dimensions of the matrix 
# (4) subset rows 10-15 and columns 3-6

#1
practice_matrix <- matrix(seq(from=2, to=500,by=2), byrow = TRUE, nrow = 25)
practice_matrix

#2
is.matrix(practice_matrix)

#3
dim(practice_matrix)

#4
practice_matrix[10:15,3:6]

```

