---
title: "Assignment 5 - dplyr and tidyr (25 pts)"
author: "CARSON CRENSHAW (cgc8gdt)"

output: html_document
---

**You will submit the html output and raw rmd file.**

```{r}
library(tidyverse)
library(ggplot2)
```

## Part 1: Let's talk about the shark in the room
Download the shark tank dataset, read over the website and columns description 
https://www.kaggle.com/datasets/thirumani/shark-tank-us-dataset
```{r}
# Load the dataset; assign the dataframe to object called 'df_shark'
df_shark <- read.csv('Shark Tank US dataset.csv')
head(df_shark)
```

## Question 1 (2 point)
Visualize the distribution of (Valuation.Requested) subsetted by (Industry)
with a violin plot.
rotate x labels 90 degrees to make them visible
```{r}
#Basic violin plot
ggplot(df_shark, aes(Industry, Valuation.Requested)) + 
  geom_violin() +
  labs(title = 'Basic Violin Plot of Valuation Requested by Industry')+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.05, hjust=1))
```

## Question 2 (2 point)
Calculate the mean, std, and count, and standard error of "Valuation Requested" 
subsetted by "Industry"
Hint: use group_by and summarize
```{r}
df_shark_summary <-df_shark%>%
  group_by(Industry) %>%
  summarize(meanval=mean(Valuation.Requested, na.rm = TRUE), 
            stdval=sd(Valuation.Requested,na.rm = TRUE),
            countval=length(Valuation.Requested), 
            seval=(stdval/sqrt(countval)))
#Using the na.rm function allows the mean and std to be calculated despite 
#NA values in the column (allows summary statistics for the children/education
#industry). There are four values in the dataset that do not have much info,
#nor do they have an industry group/valuation requested metrics (col#1). 
df_shark_summary
```

## Question 3 (2 point)
Visualize the mean (as points) and the standard error of the data (using errorbars)
rotate x labels 90 degrees to make them visible
```{r}
ggplot(df_shark_summary, aes(Industry, meanval)) + geom_point(color='orange') +
  geom_errorbar(aes(ymin=meanval-seval, ymax=meanval+seval), width=.1,color = 'orange') +
  labs(title = 'Mean and Standard Error Visualization of Valuation Requested by Industry')+
  theme_bw() +
  labs(y='Valuation.Requested Mean')+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.05, hjust=1))
```

## Question 4 (2 point)
Overlay the violin plot and the errorbar plot on the same chart
```{r}
ggplot()+ 
  geom_violin(df_shark, mapping=aes(Industry, Valuation.Requested))+ 
  geom_point(df_shark_summary, mapping=aes(Industry, y = meanval,color='orange'))+ 
  geom_errorbar(df_shark_summary, mapping=aes(x = Industry, 
                                              y = meanval, ymin = meanval - seval,
                                              ymax = meanval + seval),width = 0.5, color = "orange")+
  labs(title = 'Violin Plot of Valuation Requested by Industry with Mean and Standard Error')+
  theme_bw()+
  theme(legend.position='None')+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.05, hjust=1))
```

## Question 5 (2 point)
Do you think that using the error bars and the mean are a 
better summary of the data than the violin plot? Why or why not?
Keep it brief and direct (for grading purposes!)
```{r}
#Choosing between the errorbars/mean summary and the violin plot, the better 
#summary of the data is the violin plot. Although an audience would benefit 
#from the combined information from both the mean/errorbars and the distribution 
#of the violin plot, the advantage of the violin plot is that aside from being 
#able to show the median and quartiles it also can show the entire distribution 
#and concentration of the data. Only using the mean/standard error might skew 
#an interpretation of the data if there are outliers and does not show the 
#full extent of the data. 

#To keep it brief, the best chart would be the mean/errorbars AND the violin 
#plot, but between the two the violin plot is better than just the mean/errorbars. 
```


## Part 2: The Doggo Dataset
### Review tidyr functions to tidy a dataframe. Let's tidy this dataframe!

## Question 6 (2 point)
Import csv file and replace any values of "" with NA instead
```{r}
doggo <- read.csv('dogs.csv', na.strings=c("","NA"))
doggo
```

## Question 7 (3 point)
fill() in values for (patient), (breed), (address), and (guardian), for all rows of the dataframe obtained from Q6
(fill down same value to next value)
** Hint: fill() expects NA for fills **
```{r}
doggo1 <- doggo %>% 
  fill(patient, breed, address, guardian)
doggo1
```

## Question 8 (3 point)
Use pivot_longer() to combine (height), (length), and (width) columns to (dimension) for the dataframe obtained from Q7
```{r}
doggo2 <- doggo1 %>%
  pivot_longer(cols = c('height', 'length', 'width'), names_to = 'dimension')
doggo2
```

## Question 9 (2 point)
Split (weight) between (weight) and (weight_units) using (separate) for the dataframe obtained from Q8
```{r}
doggo3 <- doggo2 %>%
  separate(weight, c('weight', 'weight_units'), ' ')
doggo3
```

## Question 10 (2 point)
Split (visit_date) to (visit_num) and (date) using separate() function for the dataframe obtained from Q9
```{r}
doggo4 <- doggo3 %>%
  separate(visit_date, c('visit_num', 'date'), '-')
doggo4
```

## Question 11 (3 point)
Separate the dataframe from Q10 into 
* guardian_df (guardian and address), remove any repeating rows with distinct()
* dog_df (all other columns)
using the select() function
```{r}
guardian_df <- doggo4 %>%
  select('guardian', 'address') %>% distinct()
guardian_df

dog_df <- doggo4 %>%
  select(-'guardian', -'address')
dog_df

```


